%implementing document formatting:
\input{preamble.tex}
%Vectors
\renewcommand{\vec}[1]{\boldsymbol{\mathbf{#1}}}
\begin{document}
\renewcommand\chaptername{KAPITEL}
\renewcommand\contentsname{Indhold}
\renewcommand\figurename{Figur}
\renewcommand\tablename{Tabel}

%\section*{Supervisor meeting\\ \small Tuesday, 14th of February of 2017}
\section*{Sensor Fusion Research}
\subsection{Wikipedia}
\begin{itemize}
	\item https://en.wikipedia.org/wiki/Extended\_Kalman\_filter
	\item https://en.wikipedia.org/wiki/Kalman\_filter
	\item In these two pages we can find the idea of what we need.
\end{itemize}
\subsection{GPS aided attitude and Heading Reference System Using MEMS Sensors}
\begin{itemize}
	\item Contains and explanation on how to use Kalman filter for a Micro Aerial Vehicle.
	\item It only estimates three states, but the idea of it can be extracted. 
	\item It includes a good model of the gyro and an incomplete one for the accelerometer.
	\item Weird GPS model.
\end{itemize}	
\subsection{Autonomous Navigation of a Small Boat using IMU/GPS/Compass}
\begin{itemize}
	\item It considers more states, but still only in a 2-D plane. Maybe we do not need to estimate every state in the system.
	\item It includes a good model for the accelerometer and the GPS.
	\item The explanation of the kalman filter is more or less what we need in theory, but it is quite messy.
	\item The principle is: Make a model of the sensors which includes bias and noise for the accelerometers and gyro and noise for the GPS. Then make a model of the system (State Space) and include the sensor biases in the model as new states in some way.   
\end{itemize}
\subsection{INS/GPS Fusion Architectures for Unmaned Aerial Vehicles}
\begin{itemize}
	\item Same gyro and acceleration model as those in the previous articles.
	\item Good magnetometer model, I think.
	\item It ends up with a 15 states Extended Kalman Filter. It is cumbersome to read, I don't know if we can get something out of the formula as it is there.
	\item The approach is again similar to the previous articles an it seems to be the way to go.
\end{itemize}
\subsection{Adaptive Kalman Filter Based Navigation: An IMU/GPS Integration Approach}
\begin{itemize}
	\item Simpler explanation of the principle, but it is to simple as it does not include biases for sensors.
\end{itemize}
\subsection{GPS/IMU Data Fusion using Multisensor Kalman Filtering}
\begin{itemize}
	\item The equations for the filter are presented in a more detail way, but models for the sensors are not included.
	\item Good for understanding how to set up the state/process model.
\end{itemize}
\subsection{Slides: Lecture Kalman Filter}
\begin{itemize}
	\item Simple explanation with a simple example. 
	\item Good for understanding why the biases are included as states.
	\item Example with a IMU and Gyro for estimating the attitude.
\end{itemize}
\subsection{A quaternion based Unscented Kalman Filter for Orientation tracking}
\begin{itemize}
	\item Detailed, long and heavy explanation of unscented kalman filter, which is suitable for working with the system in non-linear form. 
	\item The explanation uses quaternions in the paper.
\end{itemize}
\subsection{A new Extension of the Kalman Filter to Nonlinear Systems}
\begin{itemize}
	\item General explanation of the Unscented Kalman Filter, just in case we need to look into it.
	\item I do not think we should use this.
\end{itemize}
%\subsection{Next Supervisor Meeting}
%Wednesday, 22nd of February at 13:00

\end{document}